cardinalR: Generating interesting high-dimensional data structures

### 2025-05-06 paper-cardinalR.Rmd
Added gen_linearbranches

>>>>

The `gen_linearbranches(n, p, k)` function generates a dataset of $n$ points forming $k$ approximately linear branches in a $p$-dimensional space. The core structure lies in the first two dimensions and additional dimensions carry Gaussian noise.

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches 1 and 2 are initialized with fixed slopes and intercepts. The Branch 1 is generated from $X\_1 \sim \mathcal{U}(-2, 8)\$, \$X\_2 = 0.5 \cdot X\_1 + \epsilon$. The Branch 2 is generated from $X\_1 \sim \mathcal{U}(-6, 2)\$, \$X\_2 = -0.5 \cdot X\_1 + \epsilon$. $\epsilon \sim \mathcal{U}(0, 0.5)$.

Branches 3 to $k$** are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches.

$X\_1$ values span a short segment ($x\_{start}$ to $x\_{start} + 1$),

$X\_2 = s\_i \cdot (X\_1 - x\_{start}) + y\_{start} + \epsilon$ where $s\_i$ is a sampled slope from a filtered set and $\epsilon \sim \mathcal{U}(0, 0.2)$.

This setup yields $k$ spatially distinct, noisy linear trajectories.

For $p > 2$, the remaining dimensions are pure noise:

$$
X\_j \sim \mathcal{N}(0, 0.05^2)\$ for \$j = 3, \dots, p.
$$
